---
title: "FE 800"
author: "Spencer Tirella"
date: "2022-09-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(fredr)
FRED_API_KEY="cd758d0f93b061016a32d539846b7483"
fredr_set_key(FRED_API_KEY)
```

## Data Sourcing

```{r}
FedFunds <- fredr(
  series_id = "DFF",
  observation_start = as.Date("2000-01-01"),
  observation_end = as.Date("2022-10-01"),
  frequency="m"
)
nrow(FedFunds)

TenYTwoY <- fredr(
  series_id = "T10Y2Y",
  observation_start = as.Date("2000-01-01"),
  observation_end = as.Date("2022-10-01"),
    frequency="m"
)
nrow(TenYTwoY)

NFCI <- fredr(
  series_id = "NFCI",
  observation_start = as.Date("2000-01-01"),
  observation_end = as.Date("2022-10-08"),
    frequency="m"
)
nrow(NFCI)

UnEmploy <- fredr(
  series_id = "UNRATE",
  observation_start = as.Date("1999-12-01"),
  observation_end = as.Date("2022-10-01"),
  frequency="m"
)
nrow(UnEmploy)

BusConditions <- fredr(
  series_id = "GACDISA066MSFRBNY",
  observation_start = as.Date("1999-09-01"),
  observation_end = as.Date("2022-10-01"),
  frequency="m"
)
nrow(BusConditions)

UMICH <- fredr(
  series_id = "MICH",
  observation_start = as.Date("1999-11-01"),
  observation_end = as.Date("2022-10-01"),
  frequency="m"
)
nrow(UMICH)

Recession <- fredr(
  series_id = "USRECM",
  observation_start = as.Date("1999-12-01"),
  observation_end = as.Date("2022-10-01"),
  frequency="m"
)
nrow(Recession)

LMK <- fredr(
  series_id = "FRBKCLMCIM",
  observation_start = as.Date("1999-12-01"),
  observation_end = as.Date("2022-10-01"),
  frequency="m"
)
nrow(LMK)


library("readxl")
FFFutures<-read_excel("C:/Users/spenc/Desktop/Fed Funds Futures.xlsx")
FFFutures$date <- as.Date(FFFutures$date)
FFF<-data.frame(FFFutures$date,FFFutures$`Adj Close**`)
names(FFF) <- c("date", "FFF")


allvariables <- c(FedFunds,TenYTwoY,NFCI,UnEmploy,BusConditions,UMICH,Recession, FFF)

library(dplyr)
FS1 <- merge(FedFunds,TenYTwoY, by= "date")
FS1 <- select(FS1,date,value.x,value.y)
colnames(FS1) <-c("date","FedFunds","TenYTwoY")

FS2 <- merge(FS1,NFCI, by= "date")
FS2 <- select(FS2,colnames(FS1),value)
names(FS2)[names(FS2) == "value"] <- "NFCI"

FS3 <- merge(FS2,UnEmploy, by= "date")
FS3 <- select(FS3,colnames(FS2),value)
names(FS3)[names(FS3) == "value"] <- "UnEmploy"

FS4 <- merge(FS3,Recession, by= "date")
FS4 <- select(FS4,colnames(FS3),value)
names(FS4)[names(FS4) == "value"] <- "Recession"

FS5 <- merge(FS4,BusConditions, by= "date")
FS5 <- select(FS5,colnames(FS4),value)
names(FS5)[names(FS5) == "value"] <- "BusConditions"

FS6 <- merge(FS5,UMICH, by= "date")
FS6 <- select(FS6,colnames(FS5),value)
names(FS6)[names(FS6) == "value"] <- "UMICH"

FS7 <- merge(FS6,LMK, by= "date")
FS7 <- select(FS7,colnames(FS6),value)
names(FS7)[names(FS7) == "value"] <- "LMK"

FinalSet <- merge(FS7,FFF, by= "date")

head(FinalSet)

```

## Summary Statistics
```{r}

library(ggplot2)
plot(FinalSet)

library(TSstudio)
par(mfrow = c(3, 3))
plot(FinalSet$date, FinalSet$FedFunds)
plot(FinalSet$date, FinalSet$TenYTwoY)
plot(FinalSet$date, FinalSet$NFCI)
plot(FinalSet$date, FinalSet$UnEmploy)
plot(FinalSet$date, FinalSet$BusConditions)
plot(FinalSet$date, FinalSet$UMICH)
plot(FinalSet$date, FinalSet$FFF)
plot(FinalSet$date, FinalSet$LMK)

library(GGally)
ggscatmat(FinalSet, columns = 1: ncol(FinalSet))
```


## Logistic Regression


```{r}
#Logistic Regression Full model
train <- FinalSet[1:175,]
test <- FinalSet[176:219,]

model <- glm(Recession ~ ., family=binomial,data=train)
summary(model)

test$model <- predict(model, test, type = "response")
logisticprediction <- ifelse(test$model>0.5, 1, 0)

test <-test[,c(1,6,2,3,4,5,7,8,9,10)]
Logistic_Predictions_by_date <- cbind(logisticprediction,test)
Logistic_Predictions_by_date

#Creating confusion matrix
library(caret)
logconfmatrix <- confusionMatrix(data=factor(logisticprediction), reference = factor(test$Recession))
logconfmatrix

```


# Logistc Regression Visualizations

```{r}
par(mfrow = c(2, 4))

model <- glm(Recession ~ FedFunds, family=binomial,data=FinalSet)
summary(model)
newdata <- data.frame(FedFunds=seq(min(FinalSet$FedFunds), max(FinalSet$FedFunds),len=500))
newdata$Recession = predict(model, newdata, type="response")
plot(Recession ~ FedFunds, data=FinalSet, col="steelblue")
lines(Recession ~ FedFunds, newdata, lwd=2)

model2 <- glm(Recession ~ UnEmploy, family=binomial,data=FinalSet)
summary(model2)
newdata <- data.frame(UnEmploy=seq(min(FinalSet$UnEmploy), max(FinalSet$UnEmploy),len=500))
newdata$Recession = predict(model2, newdata, type="response")
plot(Recession ~ UnEmploy, data=FinalSet, col="steelblue")
lines(Recession ~ UnEmploy, newdata, lwd=2)

model3 <- glm(Recession ~ NFCI, family=binomial,data=FinalSet)
summary(model3)
newdata <- data.frame(NFCI=seq(min(FinalSet$NFCI), max(FinalSet$NFCI),len=500))
newdata$Recession = predict(model3, newdata, type="response")
plot(Recession ~ NFCI, data=FinalSet, col="steelblue")
lines(Recession ~ NFCI, newdata, lwd=2)

model4 <- glm(Recession ~ BusConditions, family=binomial,data=FinalSet)
summary(model4)
newdata <- data.frame(BusConditions=seq(min(FinalSet$BusConditions), max(FinalSet$BusConditions),len=500))
newdata$Recession = predict(model4, newdata, type="response")
plot(Recession ~ BusConditions, data=FinalSet, col="steelblue")
lines(Recession ~ BusConditions, newdata, lwd=2)

model5 <- glm(Recession ~ UMICH, family=binomial,data=FinalSet)
summary(model5)
newdata <- data.frame(UMICH=seq(min(FinalSet$UMICH), max(FinalSet$UMICH),len=500))
newdata$Recession = predict(model5, newdata, type="response")
plot(Recession ~ UMICH, data=FinalSet, col="steelblue")
lines(Recession ~ UMICH, newdata, lwd=2)

model6 <- glm(Recession ~ FFF, family=binomial,data=FinalSet)
summary(model6)
newdata <- data.frame(FFF=seq(min(FinalSet$FFF), max(FinalSet$FFF),len=500))
newdata$Recession = predict(model6, newdata, type="response")
plot(Recession ~ FFF, data=FinalSet, col="steelblue")
lines(Recession ~ FFF, newdata, lwd=2)

model7 <- glm(Recession ~ TenYTwoY, family=binomial,data=FinalSet)
summary(model7)
newdata <- data.frame(TenYTwoY=seq(min(FinalSet$TenYTwoY), max(FinalSet$TenYTwoY),len=500))
newdata$Recession = predict(model7, newdata, type="response")
plot(Recession ~ TenYTwoY, data=FinalSet, col="steelblue")
lines(Recession ~ TenYTwoY, newdata, lwd=2)

model8 <- glm(Recession ~ LMK, family=binomial,data=FinalSet)
summary(model7)
newdata <- data.frame(LMK=seq(min(FinalSet$LMK), max(FinalSet$LMK),len=500))
newdata$Recession = predict(model8, newdata, type="response")
plot(Recession ~ LMK, data=FinalSet, col="steelblue")
lines(Recession ~ LMK, newdata, lwd=2)

```
#Logistic Regression Decision Boundary
```{r}

boundary <- function(model, data, class = NULL, predict_type = "class",
  resolution = 100, showgrid = TRUE, ...) {

  if(!is.null(class)) cl <- data[,class] else cl <- 1
  data <- data[,1:2]
  k <- length(unique(cl))

  plot(data, col = as.integer(cl)+1L, pch = as.integer(cl)+1L, ...)

  r <- sapply(data, range, na.rm = TRUE)
  xs <- seq(r[1,1], r[2,1], length.out = resolution)
  ys <- seq(r[1,2], r[2,2], length.out = resolution)
  g <- cbind(rep(xs, each=resolution), rep(ys, time = resolution))
  colnames(g) <- colnames(r)
  g <- as.data.frame(g)

  p <- predict(model, g, type = predict_type)
  if(is.list(p)) p <- p$class
  p <- as.factor(p)

  if(showgrid) points(g, col = as.integer(p)+1L, pch = ".")

  z <- matrix(as.integer(p), nrow = resolution, byrow = TRUE)
  contour(xs, ys, z, add = TRUE, drawlabels = FALSE,
    lwd = 2, levels = (1:(k-1))+.5)

  invisible(z)
}


model1 <- glm(Recession ~NFCI+BusConditions, data = FinalSet,  family=binomial(link='logit'))
class(model1) <- c("lr", class(model1))

# specify the cutoff point for prediction

x <- pp[c("NFCI", "BusConditions","Recession")]

predict.lr <- function(object, x, ...)
  predict.glm(object, x, type = "response") > .5

boundary(model1, x,class="Recession", main = "Logistic Regression Decision Boundary")


```



```{r}
# Logistic Regression 

# Likelihood ratio test
par(mfrow = c(1, 4))
Logistic.model <- glm(Recession ~ .,data=FinalSet)
summary(Logistic.model)
anova(Logistic.model)
plot(Logistic.model)


#Logistic Regression Stepwise model
par(mfrow = c(1, 4))
step.logistic.model <- step(Logistic.model, direction='both', scope=formula(Logistic.model), trace=0)
summary(step.logistic.model)
anova(step.logistic.model)
plot(step.logistic.model)

library(lmtest)
lrtest(Logistic.model, step.logistic.model) 

```


## Linear MOdel

```{r}
#Linear Regression Full model
train <- FinalSet[1:175,]
test <- FinalSet[176:219,]

linmodel <- lm(Recession ~ .,data=train)
summary(linmodel)

test$model <- predict(linmodel, test, type = "response")
linearprediction <- ifelse(test$model>0.5, 1, 0)

test <-test[,c(1,6,2,3,4,5,7,8,9,10)]
Linear_Predictions_by_date <- cbind(linearprediction,test)
Linear_Predictions_by_date

library(caret)
#Creating confusion matrix
linconfmatrix <- confusionMatrix(data=factor(linearprediction), reference = factor(test$Recession))
linconfmatrix

```

## Linear Model - Predicting Fed Funds

```{r}
#Linear Regression Full model
train <- FinalSet[1:175,]
test <- FinalSet[176:219,]

linmodel2 <- lm(FedFunds ~ .,data=train)
summary(linmodel2)
#Step
lin.step.model2 <- step(linmodel2, direction='both', scope=formula(linmodel2), trace=0)
summary(lin.step.model2)

#Points not in Step
linmodel3 <- lm(FedFunds ~ TenYTwoY+NFCI+UnEmploy+Recession+BusConditions+UMICH+LMK,data=train)
summary(linmodel3)

test$model <- predict(linmodel3, test, type = "response")
FedFundsPrediction <- test$model

Linear_Predictions_by_date2 <- cbind(FedFundsPrediction,linearprediction,test)
Linear_Predictions_by_date2

FFFnorm <- 100-test$FFF

ggplot(test, aes(date)) + 
  geom_line(aes(y = FedFunds, colour = "Actual Fed Funds")) + 
  geom_line(aes(y = FedFundsPrediction, colour = "Linear Model Feds Funds Prediction")) + 
  geom_line(aes(y = FFFnorm, colour = "Actual Fed Funds Futures"))


```


##Linear Model Decision Boundary

```{r}


model2 <- lm(Recession ~NFCI+BusConditions, data = FinalSet,  family=binomial(link='logit'))
class(model2) <- c("lr", class(model2))

# specify the cutoff point for prediction

x <- FinalSet[c("NFCI", "BusConditions","Recession")]

predict.lr <- function(object, x, ...)
  predict.lm(object, x, type = "response") > .5

boundary(model2, x,class="Recession", main = "Linear Regression Decision Boundary")
```


```{r}

par(mfrow = c(1, 4))
# Fit the full model 
lin.model <- lm(Recession ~ .,data=FinalSet)
summary(lin.model)
anova(lin.model)
plot(lin.model)

# Stepwise regression model
par(mfrow = c(1, 4))
lin.step.model <- step(lin.model, direction='both', scope=formula(lin.model), trace=0)
summary(lin.step.model)
anova(lin.step.model)
plot(lin.step.model)

# Likelihood ratio test
library(lmtest)
lrtest(lin.model, lin.step.model)

```




## Neural Network 

```{r}
library(keras)
library(mlbench)
library(dplyr)
library(magrittr)
library(neuralnet)

n <- neuralnet(Recession ~  TenYTwoY+NFCI
               +UnEmploy+BusConditions+FFF+LMK,
               data = FinalSet,
               hidden = 3,
               linear.output = F,
               act.fct = "logistic")


plot(n,col.hidden = 'darkgreen',     
col.hidden.synapse = 'darkgreen',
     show.weights = T,
     information = T,
     fill = 'lightblue')

summary(n)

```


## Neural Network Model

```{r}
## Prediction using neural network

trainnn <- FinalSet[1:175,]
testnn <- FinalSet[176:219,]

n <- neuralnet(Recession ~  TenYTwoY+NFCI +UnEmploy+BusConditions+FFF+LMK, data = trainnn,
               hidden = 3, linear.output = F,  act.fct = "logistic")

Predict=compute(n,testnn)
Predict$net.result

nnprob <- Predict$net.result
nnpred <- ifelse(nnprob>0.3, 1, 0)

testnn2 <-testnn[,c(1,6,2,3,4,5,7,8,9)]

NN_Predictions_by_date <- cbind(nnpred,testnn2)
NN_Predictions_by_date

#Creating confusion matrix
library(caret)
NNconfmatrix <- confusionMatrix(data=factor(nnpred), reference = factor(testnn$Recession))
NNconfmatrix

```


##Plots with Neural Networks

```{r}

library(zoo)
Year_mon <- as.yearmon(FinalSet$date, format = "%Y-%m-%d")
pp<-cbind(FinalSet,Year_mon)

n2 <- neuralnet(Recession ~  FedFunds+TenYTwoY+NFCI
               +UnEmploy+BusConditions+UMICH+FFF+LMK,
               data = FinalSet,
               hidden = 3,
               linear.output = F)

Z <- compute(n2,FinalSet)

breaks <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
# specify interval/bin labels
Z2 <- cut(Z$net.result,breaks=breaks)
Z3 <- cbind(Z,Z2)

library(ggplot2)

(ggplot(FinalSet,aes(x=NFCI,y=BusConditions,col=Recession))+geom_point(show.legend = F)
+geom_hline(yintercept = 0)
+geom_vline(xintercept = 0)
+geom_text(label=pp$Year_mon,check_overlap = T,show.legend = F)
+theme_bw(base_size = 15)
+ggtitle('Two Factors By Recession'))


xvalue <- NFCI
yvalue <- BusConditions

x_min <- min(xvalue$value)
x_max <- max(xvalue$value)
y_min <- min(yvalue$value) 
y_max <- max(yvalue$value)
grid <- as.data.frame(expand.grid(seq(x_min, x_max, by = (x_max-x_min)/sqrt(121)), seq(y_min, y_max, by =(y_max-y_min)/sqrt(121))))
area <- grid[1:219,]
nrow(area)
nrow(FinalSet)


(ggplot(FinalSet,aes(x=NFCI,y=BusConditions,col=Recession))
+geom_point(show.legend = F)
+geom_tile(aes(x=area$Var1,y=area$Var2,color=Z$net.result), alpha = 0.3, show.legend = T)
+scale_color_gradient("Prediction", low = "blue", high = "red")
+geom_text(label=pp$Year_mon,check_overlap = T,show.legend = F)
+geom_hline(yintercept = 0)
+geom_vline(xintercept = 0)
+ggtitle('Neural Network Decision Boundary')
)




```

## Forest Model


#Variable SIgnficance & outlier detection
```{r}
#Decision Tree
library(party)
library(rpart)
library(rpart.plot)

output.tree <- ctree(
Recession ~ TenYTwoY+NFCI +UnEmploy+BusConditions+FFF+LMK, 
data = FinalSet)

output.tree2 <- rpart(
Recession ~ TenYTwoY+NFCI +UnEmploy+BusConditions+FFF+LMK, 
data = FinalSet)

# Plot the tree.
plot(output.tree)
rpart.plot(output.tree2)

```

#Train/Test Model
```{r}
#Decision Tree 
train <- FinalSet[1:175,]
test <- FinalSet[176:219,]

  tree.model <- ctree(
  Recession ~ TenYTwoY+NFCI +UnEmploy+BusConditions+FFF+LMK, 
  data = train)
  
rfpred <- predict(tree.model, test, type = "response")
treeprediction <- ifelse(rfpred>0.5, 1, 0)

test <-test[,c(1,6,2,3,4,5,7,8,9,10)]
Tree_Predictions_by_date <- cbind(treeprediction,test)
Tree_Predictions_by_date

```

##Decision Tree Model Decision Boundary

```{r}
library(randomForest)
model3 <- randomForest(Recession ~NFCI+LMK, data = FinalSet)
x <- FinalSet[c("NFCI", "LMK","Recession")]
boundary(model3, x,class="Recession", main = "Decision Tree Decision Boundary")

(ggplot(FinalSet,aes(x=NFCI,y=LMK,col=Recession))+geom_point(show.legend = F)
+geom_hline(yintercept = -1.1)
+geom_vline(xintercept = 0.32)
+geom_text(label=pp$Year_mon,check_overlap = T,show.legend = F)
+theme_bw(base_size = 15)
+ggtitle('Decision Tree Two Factors By Recession'))

```

#Model Accuracy
```{r}
#Creating confusion matrix
library(caret)
Treeconfmatrix <- confusionMatrix(data=factor(treeprediction), reference = factor(test$Recession))
Treeconfmatrix

```
#Combined Time Series
```{r}

p1 <- ggplot(test, aes(x=date, y=logisticprediction))+  geom_bar(stat="identity")+ theme_minimal() + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
p2 <- ggplot(test, aes(x=date, y=linearprediction))+  geom_bar(stat="identity")+ theme_minimal() + 
      theme(axis.title.x = element_blank(), axis.text.x = element_blank())
p3 <- ggplot(test, aes(x=date, y=nnpred))+  geom_bar(stat="identity")+ theme_minimal() + 
      theme(axis.title.x = element_blank(), axis.text.x = element_blank())
p4 <- ggplot(test, aes(x=date, y=treeprediction))+  geom_bar(stat="identity")+ theme_minimal() + 
      theme(axis.title.x = element_blank(), axis.text.x = element_blank())
p5 <- ggplot(test, aes(x=date, y=NFCI))+ geom_line()+ theme_minimal() + 
      theme(axis.title.x = element_blank(), axis.text.x = element_blank())
p6 <- ggplot(test, aes(x=date, y=LMK))+ geom_line()+ theme_minimal() + 
      theme(axis.title.x = element_blank(), axis.text.x = element_blank())
p7 <- ggplot(test, aes(x=date, y=BusConditions))+ geom_line()+ theme_minimal() + 
      theme(axis.title.x = element_blank(),axis.text.x = element_text(angle=90))+scale_x_date(date_labels = "%m-%Y")+scale_x_date(date_breaks = "3 month") 
test

library(grid)
grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2), ggplotGrob(p3), ggplotGrob(p4), ggplotGrob(p5), ggplotGrob(p6), ggplotGrob(p7), size = "last"))

```

```{r}
logconfmatrix ##Current prediction=1  BA=0.6905
linconfmatrix ##Current prediction =0  BA = 0.7262
NNconfmatrix ##Current prediction =1  BA = 0.63095 
Treeconfmatrix ##Current prediction = 0 BA =0.7262 

```
